Social Visual Analytics: DevOps for Data Scientists

### 2015-03-20 conclusion.tex
Renamed some files

We presented a case for an environment that supports the visual
analytics process, from data acquisition and exploratory
data analysis, through development and deployment.
This approach can be considered ``DevOps for data science''
or reproducible data science.
The environment supports collaboration and communication
with capabilities for creating and sharing experimental workbooks,
version control, searching, recommending, annotating, and publishing
experiments as web sites and as reusable services.

@@

Some possible next steps are to support cross-language development,
to incorporate richer recommendation techniques, to provide fine-grained
information security, and to improve the usability of the human interface.

>>>>

We presented a case for environments that support the visual
analytics process, from data acquisition and exploratory
data analysis, through development and deployment.
The environment supports collaboration and communication
through shared experimental workbooks with version control,
searching, recommending, annotating, and publishing
experiments as web sites and as reusable services.

@@

This work prototype a step toward practical ``DevOps for data
science'' and offers reproducible experiments.
Possible next steps are to support cross-language development,
to incorporate richer recommendation techniques, to provide fine-grained
information security, and to improve the usability of the human interface.

### 2015-03-20 discussion.tex
Renamed some files

{\bf Curation.} Even without more formal group management, users are
able to curate groups of related notebooks using the simple folder
system.  This is particularly effective for bringing together
training materials.

@@

Because we developed the ideas and the prototype together,
we did not foresee some of the requirements that emerged
after people started working with the prototype.

{\bf Multiple language support.}
We did not anticipate having to back off so soon on the
assumption that programming is being done mainly in one
language (R). In retrospect, it should have been obvious
that many data science problems are already being solved
with cross-language approaches.

{\bf Versioning.}
A disadvantage is that some important aspects of the environment
are more difficult to manage in RCloud than in conventional systems.
Not only does it lack an explicit separation of the development and
deployment environments, but every version of every prototype is
shared by default. While this encourages collaboration on work
in progress, errors such as misconfiguration, mismatches between
versions of packages, and ordinary programming errors are also
easily shared and can affect production websites unintentionally.
This cannot be avoided entirely, but the power of sharing is worth
this price (as proven true with distributed systems in general).
Still, this problem needs more attention if the environment is
going to be scaled up to a wider number of users with fewer
informal channels to coordinate work. Effective sharing in RCloud
will require control over versioning.
Versioning must be managed in several places: in scripts, in the R environment
such as the installed R libraries and packages, and in the external
environment such as the operating system and even protocols spoken by
remote services. At one end of the scale, we have full control over the
versioning of scripts via git, though there is still a need for conventions
for example to name stable or working versions of scripts or versions known
to work together. The R environment is under the control of its package
manager. It is at least possible for RCloud to access information about
package versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not strong.
At the other end of the scale, there is not much reason to expect version
control for the external environment. Most programmers rely on clean living
and a conservative approach to upgrades.

On top of this, RCloud allows hackers to fork experiments to try
new ideas quickly. So far we have not done much more to address the
problem of having a lot of bits and pieces of code and data lying
around, though at least we have made it easier to get at them and
to apply algorithms and metadata to organize them.

@@

{\bf Caching.}
An important consideration is how and where to introduce caching
in RCloud's distributed computation model. Caching can dramatically
improve performance in a way that is otherwise transparent to
application program semantics \cite{Callahan:2006:VVM, Guo:2010:TPI}.
Though caching is usually implicit, in some environments, such as
Vistrails, programmers may also have some explicit control over cache
management. This may be desirable to ensure the repeatability of
computations that rely on volatile or unreliable data sources.
For example, the stock ticker use case is one for which the data
might temporarily become unavailable, so caching could improve
reliability and consistency of results.  Alternatively, in a use case
involving text analysis through a topic modeling technique such as
Latent Dirichlet Allocation (LDA) that is relatively expensive,
we might want caching to improve performance. Currently we would
program this in RCloud by explicitly saving the analysis in a
persistent database, but it would be desirable to implement this
in a more general purpose associative cache. We envision cache
management being implemented in a new middle layer not in RCloud's
architecture today.

@@

Our back-end environment assumes a high degree of trust between users.
Access control for information security is delegated to the host
operating system and web server. In practice, most operating
systems and web servers rely either on coarse permission models,
which tend to be ineffective, or on detailed access control
lists, which tend to be cumbersome. 
Information security in RCloud should be re-examined, but more
sophisticated approaches such as information flow analysis are
still an active research area, and many subtle problems remain
\cite{Moore:2011:SAF}. We are hopeful that having a richer
environment for collaboration and information sharing will encourage
practical applications for new approaches to information security
in visual analytics.

>>>>

{\bf Curation.} Even without formal group management, users are
able to curate groups of related notebooks using the folder
system. This is particularly effective for collecting and
distributing training materials.

@@

Because we developed these ideas while creating a prototype,
we did not foresee some of the requirements that emerged after
people started working with it.

{\bf Multiple language support.}
We did not anticipate having to back off so soon on the assumption that
programming is done mainly in R.  In retrospect, it is clear that many
data science problems are solved with cross-language approaches
(for example, scikit-learn in Python, weka in Java).

{\bf Versioning.}
Some important aspects of the environment are more difficult to manage
in RCloud than in conventional systems.  RCloud does not explicitly
separate the development and deployment environments. Actually, every
version of every prototype is shared by default. Although this encourages
collaboration on work in progress, it also quickly exposes errors such as
misconfiguration, mismatches between versions of packages, and programming
errors that can affect production websites unintentionally.
This cannot be avoided entirely, but the power of sharing is often worth
this price (as proven true with distributed systems in general).
Still, effective control over versioning when sharing 
problem needs more attention if the environment is
going to be scaled up to a wider number of users with fewer
informal channels to coordinate work.
Versioning must be managed in several places: in scripts, in the R environment
such as the installed R libraries and packages, and in the external
environment such as the operating system and even protocols spoken by
remote services. At one end of the scale, we have full control over the
versioning of scripts via git, along with conventions to name stable or
working versions of scripts or versions known to work together.
The R environment is under the control of its package manager.
It is at least possible for RCloud to access information about
package versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not strong.
At the other end of the scale, there is not much reason to expect version
control for the external environment. Most programmers rely on clean living
and a conservative approach to upgrades.

On top of this, RCloud allows hackers to fork experiments to try
new ideas quickly. So far we have not done much more to address the
problem of having a lot of bits and pieces of code and data lying
around, though we have provided a structure in which it should be
easier to find them and to apply algorithms and metadata to organize them.
\stephen{Why should the reader believe that?}

@@

{\bf Caching.}
An important consideration is how and where to introduce caching
in RCloud's distributed computation model. Caching can dramatically
improve performance in a way that is otherwise transparent to
application program semantics \cite{Callahan:2006:VVM, Guo:2010:TPI}.
Though caching is usually implicit, in some environments, such as
Vistrails, programmers may also have some explicit control over cache
management. This may be desirable to ensure the repeatability of
computations that rely on volatile or unreliable data sources.
For example, the stock ticker use case is one for which the data
might temporarily become unavailable, so caching could improve
reliability and consistency of results.  Alternatively, in situations
involving relatively expensive analysis (for example, analysis of large
text corpora, clustering of multivariate time series, deep learning, etc.)
caching may be essential to provide adequate performance. Currently we would
program this in RCloud by explicitly saving the analysis in a persistent database,
but it would be desirable to implement this in a general purpose associative
cache, not in application-dependent libraries.  We envision cache management
being implemented in a new middle layer not in RCloud's architecture today.

@@

Our back-end environment assumes a high degree of trust between users.
Access control for information security is delegated to the host
operating system and web server. In practice, most operating
systems and web servers rely either on coarse permission models,
which tend to be ineffective, or on detailed access control
lists, which tend to be cumbersome. 
Information security in RCloud should be improved, but more
sophisticated approaches (such as information flow analysis,
and query languages that ensure differential privacy) are
active research topics, and many difficult problems remain
\cite{Moore:2011:SAF}. We are hopeful that having a richer
environment for collaboration and information sharing will encourage
new approaches to information security in visual analytics.
